### 2025-11-18 basic-statistics.qmd
20251111 - the end of average
20251118 - squared semi-partial correlation

That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
One issue with the mean is that it is sensitive to extreme (outlying) values.
For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} petersenlab::apa(exampleValues_mean, decimals = 2)`.

@@

@@

| Effect Size Index                                   | Small       | Medium      | Large       |
|:----------------------------------------------------|:------------|:------------|:------------|
| Cohen's $d$                                         | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
| Standardized regression coefficient (beta; $\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Correlation coefficient ($r$)                       | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Cohen's $\omega$                                    | $\ge .10$   | $\ge .30$   | $\ge .50$   |
| Cohen's $f$                                         | $\ge .10$   | $\ge .25$   | $\ge .40$   |
| Cohen's $f^2$                                       | $\ge .01$   | $\ge .06$   | $\ge .16$   |
| Coefficient of determination ($R^2$)                | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Eta squared ($\eta^2$)                              | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Partial eta squared ($\eta_p^2$)                    | $\ge .01$   | $\ge .06$   | $\ge .14$   |

>>>>

That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
One issue with the mean is that it is sensitive to extreme (outlying) values.
For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} petersenlab::apa(exampleValues_mean, decimals = 2)`.
Another issue with the mean is that it may not be the same as any individual's value [@Rose2016]—that is, it may be unrepresentative of all of the individuals whose values comprise the average.

@@

#### Squared Semi-Partial Correlation ($sr^2$) {#sec-squaredSemiPartialCorrelation}

Squared semi-partial correlations are described in @sec-multipleRegressionExplainedVariancePredictors.

@@

| Effect Size Index                                   | Small       | Medium      | Large       |
|:----------------------------------------------------|:------------|:------------|:------------|
| Cohen's $d$                                         | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
| Standardized regression coefficient (beta; $\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Correlation coefficient ($r$)                       | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Cohen's $\omega$                                    | $\ge .10$   | $\ge .30$   | $\ge .50$   |
| Cohen's $f$                                         | $\ge .10$   | $\ge .25$   | $\ge .40$   |
| Cohen's $f^2$                                       | $\ge .01$   | $\ge .06$   | $\ge .16$   |
| Coefficient of determination ($R^2$)                | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Squared semi-partial correlation ($sr^2$)           | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Eta squared ($\eta^2$)                              | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Partial eta squared ($\eta_p^2$)                    | $\ge .01$   | $\ge .06$   | $\ge .14$   |

### 2025-11-18 getting-started.qmd
20251118 - squared semi-partial correlation

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr",
     "viridis","viridisLite","msir","plotly","webshot2","quantmod","fPortfolio",
     "NMOF","nFactors","xts","zoo","forecast","parallelly","doParallel",
     "missRanger","ggridges","powerjoin","bestNormalize","LongituRF","gpboost",
     "mgcv","rms","car","lavaan","lavaanPlot","lavaangui","mice","miceadds",
     "interactions","robustbase","ordinal","MASS","data.table","future",
     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

>>>>

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr",
     "viridis","viridisLite","msir","plotly","webshot2","quantmod","fPortfolio",
     "NMOF","nFactors","xts","zoo","forecast","parallelly","doParallel",
     "missRanger","ggridges","powerjoin","bestNormalize","LongituRF","gpboost",
     "mgcv","rms","car","lavaan","lavaanPlot","lavaangui","mice","miceadds",
     "interactions","robustbase","ordinal","MASS","data.table","future",
     "future.apply","SimDesign","domir","GGally","Rglpk","TTR","rockchalk"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

### 2025-11-18 index.qmd
20251118 - squared semi-partial correlation

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR","rockchalk",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

### 2025-11-18 multiple-regression.qmd
20251110 - model as simplification of reality
20251114 - absolute value of total effect
20251118 - squared semi-partial correlation

```{r}
library("petersenlab")
library("rms")
library("car")
library("bestNormalize")
library("lme4")
library("performance")
library("lavaan")
library("mice")
library("miceadds")
library("interactions")
library("brms")
library("parallelly")
library("robustbase")
library("ordinal")
library("MASS")
library("broom")
library("effectsize")
library("tidymodels")
library("tidyverse")
library("knitr")
```

@@

All statistical analyses follow the same basic structure, as in @eq-analysisBasicStructure [@Brauer2018]:

@@

@@

@@

@@

@@

@@

@@

>>>>

```{r}
library("petersenlab")
library("rms")
library("car")
library("bestNormalize")
library("lme4")
library("performance")
library("lavaan")
library("mice")
library("miceadds")
library("interactions")
library("brms")
library("parallelly")
library("robustbase")
library("ordinal")
library("MASS")
library("broom")
library("rockchalk")
library("effectsize")
library("tidymodels")
library("tidyverse")
library("knitr")
```

@@

In multiple regression, we estimate models.
A model is a simplification of reality.
All statistical analyses follow the same basic structure, as in @eq-analysisBasicStructure [@Brauer2018]:

@@

### Unstandardized Regression Coefficient {#sec-multipleRegressionComponentsUnstandardizedCoefficient}

@@

### Standardized Regression Coefficient {#sec-multipleRegressionComponentsStandardizedCoefficient}

@@

### Standard Error {#sec-multipleRegressionComponentsStandardError}

@@

### Confidence Interval {#sec-multipleRegressionComponentsConfidenceInterval}

@@

## How Much Variance Each Predictor Explains {#sec-multipleRegressionExplainedVariancePredictors}

Another estimate of [effect size](#sec-practicalSignificance) in addition to [standardized regression coefficients](#sec-multipleRegressionComponentsStandardizedCoefficient) is the squared semi-partial correlation ($sr^2$).
The squared semi-partial correlation reflects the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is uniquely explained by each [predictor variable](#sec-correlationalStudy).
The total model $R^2$ minus the sum of all the squared semi-partial correlations reflects the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the common variance among the [predictor variables](#sec-correlationalStudy).

@@

### Squared Semi-Partial Correlation ($sr^2$) {#sec-multipleRegressionSquaredSemiPartialCorrelation}

To estimate the [squared semi-partial correlation](#sec-multipleRegressionExplainedVariancePredictors) ($sr^2$), we use the `rockchalk::getDeltaRsquare()` function from the `rockchalk` package [@R-rockchalk].

```{r}
rockchalk::getDeltaRsquare(linearRegressionModel)
```

@@

  # total effect (c)
  total := direct + indirect
  totalAbs := abs(direct) + abs(indirect)

  # proportion mediated
  Pm := abs(indirect) / totalAbs
'
```
